---
title: "Cleaning"
author: "Michael Topper"
date: "3/4/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(fastDummies)
df = board_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-03-12/board_games.csv")

theme_set(theme_bw())
```


Exploratory Data Analysis
```{r}

```

Finding the top types of games that are created. 
```{r}
catvars = df %>% 
  select(game_id, name, family, expansion, category, artist, designer, average_rating) %>% 
  gather(type, value, -game_id, -name) %>% 
  filter(!is.na(value)) %>% 
  separate_rows(value, sep = ',') %>% 
  arrange(game_id)

catcounts = catvars %>% 
  count(type, value,  sort = T)

catcounts %>% 
  filter(type == 'category') %>% 
  mutate(value = fct_reorder(value, n)) %>% 
  top_n(20, n) %>% 
  ggplot(aes(value, n)) +
  geom_col(fill = "blue") +
  coord_flip() +
  ylab("Total") +
  xlab("Board Game Type") +
  labs(title = "Most Popular Types of Board Games")
```

Notice that maxplaytime and playtime are exactly the same!

```{r}
identical(df$max_playtime, df$playing_time)
```

Histogram of the average rating
```{r}
df %>% 
  ggplot(aes(x = average_rating)) +
  geom_histogram(fill = "blue") +
  geom_vline(aes(xintercept = quantile(average_rating, .80), color = "red")) +
  labs(color = "80th Percentile", title = "Distribution of Average Ratings Across Games") +
  xlab("Average Rating") +
  ylab("Count")
```




Making a bar graph of decade 
```{r}
df %>% mutate(decade = 10 * (year_published %/% 10)) %>% 
  group_by(decade) %>% summarise(min_average_players = mean(min_players), avgdecaderating = mean(average_rating)) %>% 
  ggplot(aes(x = decade, y = avgdecaderating)) + 
  geom_point() + 
  geom_smooth(method = 'lm') +
  labs(title = "Average Decade Rating Over Time") +
  xlab("Decade") +
  ylab("Average Rating in the Decade") +
  theme(plot.title = element_text(hjust = 0.5)) 


```

Designers by decade.
```{r}
df %>% mutate(decade = 10 * (year_published %/% 10)) %>% 
  filter(decade > 1950) %>% 
  filter(designer != '(Uncredited)') %>% 
  group_by(decade) %>% 
  count(designer, sort = T) %>%  
  filter(!is.na(designer)) %>% 
  mutate(designer = fct_reorder(designer, n, .desc = F)) %>% 
  top_n(5, n) %>% 
  ggplot(aes(x = designer, y = n, fill = as.factor(decade) )) +
  geom_col() +
  facet_wrap(~decade) +
  coord_flip() +
  labs(fill = "Decade", title = "Top 5 Designers by Decade") +
  xlab("Designer") +
  ylab("Count")

```


Most popular play styles bar graph
```{r}

df %>% 
  mutate(decade = 10 * (year_published %/% 10)) %>% 
  separate(mechanic, into = 'mechanic1' , sep = ",") %>% 
  filter(!is.na(mechanic1)) %>% 
  count(mechanic1, sort = T) %>%
  mutate(mechanic1 = fct_reorder(mechanic1, n, .desc = F)) %>% 
  top_n(20, n) %>% 
  ggplot(aes(x = mechanic1, y= n)) +
  geom_col(fill = "blue") +
  coord_flip() +
  xlab("Game Type") +
  ylab("Count") +
  labs(title = "Most Popular Play Styles for Board Games")
  
```


# Danny Cleaning

Let's just use the first mechanic the main mechanic and then count the number of expansions each game has
```{r}
danny.dat<-df %>% 
  separate(.,mechanic, "mechanic1",sep=",",extra="drop") %>% #Just grabbing the first mechanic for our purposes 
  mutate(expansion1=ifelse(is.na(expansion),0,str_count(df$expansion,",")))

#Michael is now going to do the same thing as you did with mechanic but do it with family instead
danny.dat = danny.dat %>% 
  separate(family, into = "family1", sep = ":", extra = "drop")

danny.dat %>% 
  ggplot(., aes(I(log(expansion1+1)), average_rating))+
  geom_point()+
  geom_smooth()
```
Looks like more is better here. The relationship is a little funky but overall increasing. The relationship looks best as a log relationship. So I'm going to change expansion1 to that.

```{r}
danny.dat.use<-danny.dat %>% 
  mutate(expansion1=log(expansion1+1),
         decade = 10 * (year_published %/% 10),
         above7 = as.numeric(ifelse(average_rating >= 7, 1 , 0))
         ) %>% 
  select(c(above7,max_playtime,min_age,min_players,mechanic1,users_rated,expansion1,decade)) %>% 
  #dummy_cols(., remove_selected_columns = TRUE) %>%  #Couldn't get mechanic1 to register so have to do a trick
  mutate(above7=as.factor(above7)) %>% 
  drop_na()

```

#Training and Testing

```{r}
train<-sample(seq(1,dim(danny.dat.use)[1]), .75*dim(danny.dat.use)[1], replace = FALSE)

train.dat<-danny.dat.use[train,]
test.dat<-danny.dat.use[-train,]

```

#Logistic Regression
```{r}
y.hat.log<-as.numeric(predict(glm(above7~.,data = train.dat, family = binomial(link="logit")), type="response")>.5)

length(y.hat.log)
length(train.dat$above7)

table(train.dat$above7,y.hat.log)

mean(y.hat.log ==train.dat$above7)

```

