---
title: "Exploratory Analysis- Michael"
author: "Michael Topper"
date: "2/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(ipumsr)
library(dplyr)
library(stringr)
######################UPLOADING DATA###############################
# NOTE: To load data, you must download both the extract's data and the DDI
# and also set the working directory to the folder with these files (or change the path below).
if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")
ddi <- read_ipums_ddi("cps_00004.xml")
data <- read_ipums_micro(ddi)
```

Poking around 
```{r}
data %>% 
  View()
```


Here is the definition of all the marital statuses:
```{r}
ddi[["var_info"]][[4]][[14]]
```

Hence, 4 is our indicator for whether someone is divorced or not. We won't be interested in those who were never married/single. If we are interested in determinants of divorce, it seems weird to be using people who were never married.

```{r}
new.data<-data %>% 
  filter(MARST!=6)
```


Let's look at some pretty graphs. 

```{r}
data %>% filter(MARST == 4) %>% 
  select(MARST, AGE, EDUC, HISPAN, ASIAN, SEX) %>% 
  View()

data %>% distinct(OCC)
data %>% filter(MARST == 4, INCTOT != 99999999) %>% 
  ggplot(aes(AGE, INCTOT)) +
  geom_point()
data %>% 
  filter(INCTOT == 99999999) %>% 
  count()
```


Changing the education variable to something we can work with. 
```{r}
new.data = data %>% 
  mutate(under.highschool = ifelse(EDUC %in% c(10,11,12,13,14,20,21,22,30,31,32,40,50,60,70,71,72), 1, 0)) %>% 
  mutate(highschool.only = ifelse(EDUC %in% c(73), 1, 0)) %>% 
  mutate(some.college = ifelse(EDUC %in% c(80,81,90,91,92,100,110),1,0)) %>% 
  mutate(bachelors = ifelse(EDUC %in% c(111), 1 , 0) ) %>% 
  mutate(graduate = ifelse(EDUC %in% c(120,121,122,123,124,125), 1 , 0)) 


new.data$education = 0
new.data$education[new.data$highschool.only == 1] = 1
new.data$education[new.data$some.college == 1] = 2
new.data$education[new.data$bachelors == 1] = 3
new.data$education[new.data$graduate == 1] = 4

new.data %>% select(education, RACE, AGE) %>% View()

```
 
The dataset reports many different races:
```{r}
kable((ddi[["var_info"]])[[4]][[13]], col.names = c("",""), caption = "Race Variables")
```

```{r}
kable(table(data$RACE), caption = "Frequency of Races")
```
It's concerning some of the race breakdowns only appear twice. We make the simplifying assumption to define every person with more than one race identified as multi-racial.

In order to convert variables, first I will make some dataframes to make the process less painful.
```{r}
birthplace<-(ddi[["var_info"]])[[4]][[20]]
occ.last.year<-(ddi[["var_info"]])[[4]][[23]]
```


```{r}
new.data<-new.data %>% 
  mutate(race=ifelse(RACE<701,RACE,1000)) %>% 
  mutate(race=recode(race,`100`="white",`200`="black",`300`="American Indiance/Aleut/Eskimo",`650`="Asian or Pacific Islander",`651`="Asian",`652`="Hawaiian",`700`="Other",`1000`="Multi-Racial"),
         birthplace=factor(BPL, levels = birthplace$val, labels = birthplace$lbl),
         )
```

Let's look at the barplot now:

```{r}
new.data %>% 
  group_by(race) %>% 
  summarise(n=n()) %>% 
  mutate(percent=n/sum(n)) %>% 
  ggplot(., aes(race,percent))+
  geom_bar(stat = "identity")+
  ggtitle("Percent of Races")
```
The sample is disproportionately white, but the proportions are slightly better now.

